A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-08-24 tex/paper.tex
new Gl run
new PHOENIX run
Working parallel prototype

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \figcaption{({\it top}) The line-of-sight velocity distribution function, $\varphi_v$, and its 
decomposition into broadening kernels.  The instrumental kernel ({\it dotted}) is treated as 
a Gaussian, the rotation kernel ({\it dashed}) is a Gaussian-like function of the projected 
rotational velocity, and the Doppler kernel ({\it solid}) is a $\delta$-function that introduces 
the radial velocity.  In this specific case, $\sigma_v = 2.9$\,km s$^{-1}$, $v\sin{i} = 5$\,km 
s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{sec:examples}.x.
({\it bottom}) A segment of a raw, full-resolution model spectrum and its post-processed equivalent 
after convolution by $\varphi_v$ and re-sampling at the coarser resolution of the detector pixels. 
\label{fig:broadening}}
\end{center}
\end{figure}

@@

\begin{figure}[!b]
\begin{center}
  \includegraphics{figs/chebyshev.pdf}
  \figcaption{
The spectrum at the overlap of two echelle orders (23 \& 24). \textbf{Panel 1}: The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders. \textbf{Panel 2}: To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model. \textbf{Panels 3 \& 4}: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4). This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
\label{fig:chebyshev}}
\end{center}
\end{figure}

@@

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
  parameters drawn from the posterior distribution, and the residuals that
  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
  to show the mildly covariant structure produced by slight mismatch between
  the data and model spectra.   \textbf{Right} The autocorrelation of the
residual sequence shown at left. Notice that there is significant correlation
for offsets of $\lesssim 5$ pixels.}
\label{fig:class0}
\end{center}
\end{figure}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/matrix_compilation.pdf}
\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals. \textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region. Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix. \textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column. At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region. \textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum. \textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA. \textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals. Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines. The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit. Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not. All parameters of the covariance kernels are determined self-consistently along with the stellar parameters. }
\label{fig:matrix}
\end{center}
\end{figure*}

@@

Local kernels are instantiated by the following procedure:  An ``average''
residual spectrum is created by combining ~500 spectra residuals stored from a
burn-in period using only global covariance kernels (thinned to account for
autocorrelation of the Markov chain). Then, using an iterative thresholding
criterion similar in procedure to ``sigma clipping,''\footnote{Although the
  procedure of identifying outliers is similar to sigma clipping, the inference
  on the stellar parameters is emphatically \emph{not} the same as sigma
clipping, since we now determine the weights of these spectral outliers inside
of a probabilistic framework, rather than directly setting their weights to
zero as in sigma clipping.} we identify the largest outlier pixel not already
covered by a local kernel, instantiate a local kernel at its location, and
repeat until there are no pixels above a certain threshold left.\footnote{There
  is no practical reason to delete local kernels once instantiated.  If the
  parameters have changed such that a given local kernel is no longer required,
  that kernel amplitude will be driven towards zero and represent a negligible
  contribution to $\vC$; in effect, the model will act as if the kernel were
deleted automatically.} Alternative local kernel instantiation schemes, such as
re-evaluating the location of kernels with each Gibbs sampler iteration, have
yielded similar results to the thresholding approach. However, this approach
most consistently converges and only has a modest computational overhead. We
choose four times the standard deviation of the average residual spectrum as
our threshold. Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the
instantiation of more kernels, which in turn slightly reduces the amplitude of
the global kernel. Taken to the logical conclusion, if the threshold was set
low enough then there would be a local kernel for each spectral line and no
global kernel would be required. We find that the final stellar posterior is
relatively insensitive to the choice of threshold level as long as it is set
low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
Once all of the local kernels have been instantiated, the Gibbs sampler is run
for another period of burn-in.

@@

In this section, we aim to demonstrate the functionality and highlight the flexibility and utility 
of the probabilistic inference methodology described above.  To do that, we illustrate how the 
modeling framework operates for two representative examples.  The first is an elaboration of the 
example shown throughout Section \ref{sec:method}, using a high resolution optical 
($\sim$5000-5400\,\AA) spectrum of the F5 dwarf main sequence star (and transiting exoplanet host) 
WASP-14 \citep{joshi09,torres12}.  This spectrum is interpreted in the context of two different 
synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
$\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
and the most recent incarnation of the {\sc Phoenix} library \citep{husser13}.  The second 
example uses a medium resolution near-infrared ($\sim$2.0-2.4\,$\mu$m) spectrum of the M5 dwarf 
star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
\citep{cushing05,rayner09}.  In this case we adopt the standard {\sc BT-Settl} modification of the 
{\sc Phoenix} library \citep{allard12}.

@@

A high resolution ($R\approx44,000$) optical spectrum of WASP-14 was obtained on 2009 June 14 
using the Tillinghast Reflector Echelle Spectrograph \citep[TRES;][]{furesz08} on the Fred Lawrence 
Whipple Observatory 1.5\,m telescope.  TRES delivers a 51-order echelle spectrum covering the 
full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
standard techniques in the TRES pipeline (cf.,~\citealt{mink11}; see \citealt{torres12} for more 
specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.

\citet{torres12} originally acquired these data in an effort to measure the WASP-14 stellar 
properties, and thereby better constrain the parameters of the transiting exoplanet it hosts.  To 
that end, they employed the {\tt MOOG} and {\tt SPC} approaches outlined in Section 
\ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
free parameter or assumed to be fixed (to $\log g = 4.29$), based on an inference of the mean 
stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
evolution models in the H-R diagram \citep{joshi09}.  Only parameters inferences from the latter 
assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
synthesize a suite of individual \ion{Fe}{1} and \ion{Fe}{2} lines in the $\sim$4300--6750\AA\ 
region, whose equivalent widths were compared with the TRES measurements in a simple least-squares 
sense.  This approach suggests that $T_{\rm eff} = 6150\pm75$\,K and $[{\rm Fe/H}] = 
-0.34\pm0.10$\,dex.  The {\tt SPC} analysis relied on a direct comparison between three TRES 
spectral orders covering $\sim$5100-5400\,\AA\ and synthetic spectra from the {\sc CfA/Kurucz} 
library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
i = 3.9\pm0.5$\,km s$^{-1}$.  \hili{It is worth noting that these parameter inferences represent an 
average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
the systematic spread between the best-fit parameter values.} \comm{This is something we need to 
discuss more with Dave and Willie.}

Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
it is most relevant to compare the \citeauthor{torres12}~{\tt SPC} result with our framework.  
Using the same TRES spectral orders (for only a single WASP-14 observation, as noted above), the 
{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer 
parameters that are entirely consistent with the {\tt SPC} approach: $T_{\rm eff} = 6517\pm13$\,K, 
$[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and $v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~X displays 
some relevant marginal projections of the posterior-space, and Table~X compiles our estimates of 
the peaks of the marginal posteriors and associated confidence intervals for all parameters 
employed in the modeling.  Figure~Y demonstrates the overall quality of this approach in 
reproducing the WASP-14 spectrum and, most significantly, the structure of the associated 
residuals.  Figure~Z compares the inferred parameters with the various determinations made by 
\citet{torres12}.

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_Kurucz_logg.pdf}
\caption{Kurucz WASP-14. One, two, and three sigma contours formed from 200 random draws from the covariance matrix.}
\label{fig:residuals_Kurucz}
\end{center}
\end{figure*}

\begin{figure}[!htb]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler. These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$). Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
}
\label{fig:stellar_posterior}
\end{center}
\end{figure}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6517.1 & 12.5 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.620 & 0.016 $\kms$ \\
$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
$\cc{22}{1}$ & -0.0142 & 0.0015\\
$\cc{22}{2}$ & -0.0145 & 0.0017\\
$\cc{22}{3}$ & -0.0031 & 0.0011\\
\\
$b_{22}$ & 1.04 & 0.02 \\
$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.020 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
$l_{22}$ & 7.54 & 0.77 $\kms$ \\

\cutinhead{Order 23 nuisance parameters}
$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
$\cc{23}{1}$ & -0.0133 & 0.0016 \\
$\cc{23}{2}$ & -0.0247 & 0.0014 \\
$\cc{23}{3}$ &  0.0012 &  0.0013\\
\\
$b_{23}$ & 1.01 & 0.02 \\
$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.015 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
$l_{23}$ & 6.52 & 0.50 \\

\cutinhead{Order 24 nuisance parameters}
$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00} \\
$\cc{24}{1}$ & -0.0103 & 0.0016 \\
$\cc{24}{2}$ & -0.0112 & 0.0013\\
$\cc{24}{3}$ &  0.0002 & 0.0012\\
\\
$b_{24}$ & 1.02 & 0.02 \\
$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
$l_{24}$ & 7.33 & 0.59 $\kms$ \\
\enddata
\end{deluxetable}

@@

\comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
want to see Table 2 in this paper, or any reference to specific model or test identifications.  It 
is too confusing, and it obscures the key points we want to make.  I appreciate your desire to 
demonstrate the complexity of the model sequentially, but it does not add anything useful/tangible 
here: it is already done very well in Section 2.  I feel like up to here we have 
established ``reproducibility" well.  The remainder of this section should be solely about 
systematics.  First, I want a paragraph breaking down the origins of the uncertainties (see the 
paragraph I commented out below...I don't like it, but it should give you an idea about what I 
mean.  If I interpret your Google tables correctly, you're getting about half the uncertainty from 
Poisson noise, and the other half from the global covariance kernel.  This is where you need to be 
able to defend the seemingly small uncertainties you're finding, even with a non-trivial covariance 
matrix.  Is it really true that a model with 0.03\,dex lower metallicity cannot describe the data 
as well?  Why?  [This is not to say I don't believe it, but it {\it will} raise questions...]}

Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
range that no local covariance kernels need to be instantiated; in this specific case we use only a 
global covariance kernel.}  To break down their respective contributions, we performed the 
same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
are generated from the model; in this paradigm, the only way to introduce deviation is through 
white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
posteriors.

\comm{Next, you want to address the observational systematics issue.  Willie (and basically 
everyone else) add empirically-estimated systematic terms in quadrature that make some pretty 
unjustified assumptions: namely that they are Gaussian and independent in each parameter.  If I 
understant Willie's paper correctly, there are 4 TRES observations of WASP-14.  If that is true, I 
think we need to fit each of them independently.  I'd like to know what ``systematic" scatter you 
would derive if you wanted to employ the standard approach.  My gut tells me that it's important to 
simultaneously model all 4 spectra in orders 22-24, but if you think that's not 
computationally tractable, we can discuss more.}
\ctext{It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}). Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$. Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.}

>>>>

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \figcaption{({\it top}) The line-of-sight velocity distribution function, $\varphi_v$, and its 
decomposition into broadening kernels.  The instrumental kernel ({\it dotted}) is treated as 
a Gaussian, the rotation kernel ({\it dashed}) is a Gaussian-like function of the projected 
rotational velocity, and the Doppler kernel ({\it solid}) is a $\delta$-function that introduces 
the radial velocity.  In this specific case, $\sigma_v = 2.9$\,km s$^{-1}$, $v\sin{i} = 5$\,km 
s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{subsec:wasp}.
({\it bottom}) A segment of a raw, full-resolution model spectrum and its post-processed equivalent 
after convolution by $\varphi_v$ and re-sampling at the coarser resolution of the detector pixels. 
\label{fig:broadening}}
\end{center}
\end{figure}

@@

\begin{figure}[!b]
\begin{center}
  \includegraphics{figs/chebyshev.pdf}
  \figcaption{
    The spectrum at the overlap of two echelle orders (23 \& 24). ({\it panel 1}) The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders. ({\it panel 2}) To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model. In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be ({\it panel 3}), but instead we multiply the model by the polynomials to recover the original discrepancy between orders ( {\it panel 4}). This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
\label{fig:chebyshev}}
\end{center}
\end{figure}

@@

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \caption{ ({\it top}) A typical model fit using 
  parameters drawn from the posterior distribution and the residuals that
  result. ({\it middle}) The low-amplitude, mildly covariant residuals in the gray band enlarged
  to show the mildly covariant structure produced by slight mismatch between
  the data and model spectra. ({\it bottom}) The autocorrelation of the
residual sequence shown at left. Notice that there is significant correlation
for offsets of $\lesssim 8$ pixels.}
\label{fig:class0}
\end{center}
\end{figure}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/matrix_compilation.pdf}
\caption{ ({\it top}) A typical model fit and residual spectrum (zoomed). An illustrative region is shaded in grey, showing the scale and structure of the residuals as an example for the following panels. ({ \it left column}) A small inset of the covariance matrix corresponding to the gray region. Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix. ({\it right column}) Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column. At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region. {\it first row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) poorly reproduces both the scale and structure of a typical residual spectrum. ({\it second row}) The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA. ({\it third row}) Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals. Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines. The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit. Stated differently, random draws from the likelihood function now behave like the residuals, whereas white noise cannot. All parameters of the covariance kernels are determined self-consistently along with the stellar parameters. }
\label{fig:matrix}
\end{center}
\end{figure*}

@@

Local kernels are instantiated by the following procedure:  an ``average''
residual spectrum is created by combining ~500 spectra residuals stored from a
burn-in period using only global covariance kernels (prior to storage the
Markov chain is thinned to account for autocorrelation of the posterior
samples). Then, using an iterative thresholding criterion similar in procedure
to ``sigma clipping,''\footnote{Although the procedure of identifying outliers
  is similar to sigma clipping, the inference on the stellar parameters is
  emphatically \emph{not} the same as sigma clipping, since we now determine
  the weights of these spectral outliers inside of a probabilistic framework,
  rather than directly setting their weights to zero as in sigma clipping.} we
  identify the largest outlier pixel not already covered by a local kernel,
  instantiate a local kernel at its location, and repeat until no residual
  pixels above a certain threshold remain. Alternative local kernel
  instantiation schemes, such as re-evaluating the location of kernels with
  each Gibbs sampler iteration, yield similar results to the thresholding
  approach. However, we found that this thresholding approach most consistently
  converged with the least computational overhead. We explored the behavior of
  setting different threshold levels, and decided to use four times the
  standard deviation of the average residual spectrum. Lower thresholds (e.g.,
  3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which
  in turn slightly reduces the amplitude of the global kernel. Taken to the
  logical conclusion, if the threshold was set low enough then there would be a
  local kernel for each spectral line and no global kernel would be required.
  We find that the final stellar posterior is relatively insensitive to the
  choice of threshold level as long as it is set low enough to capture the most
  egregious spectral outliers (e.g., 4$\sigma$).  Once all of the local kernels
  have been instantiated, the Gibbs sampler is run for another period of
  burn-in.\footnote{There is no practical reason to delete local kernels once
    instantiated.  If the parameters have changed such that a given local
    kernel is no longer required, that kernel amplitude will be driven towards
    zero and represent a negligible contribution to $\vC$; in effect, the model
    will act as if the kernel were deleted automatically.}

@@

In this section, we aim to demonstrate the functionality and highlight the flexibility and utility 
of the probabilistic inference methodology described above.  To do that, we illustrate how the 
modeling framework operates for two representative examples.  The first is an elaboration of the 
example shown throughout Section \ref{sec:method}, using a high resolution optical 
($\sim$5000-5400\,\AA) spectrum of the F5 dwarf main sequence star (and transiting exoplanet host) 
WASP-14 \citep{joshi09,torres12}.  This spectrum is interpreted in the context of two different 
synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
$\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
and the most recent incarnation of the {\sc Phoenix} library \citep{husser13}.  The second 
example uses a medium resolution near-infrared ($\sim$2.0-2.4\,$\mu$m) spectrum of the M5 dwarf 
star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
\citep{cushing05,rayner09}.  In this case we also use the {\sc Phoenix} library, since it forms a complete grid for late spectral types.

@@

A high resolution ($R\approx44,000$) optical spectrum of WASP-14 was obtained on 2009 June 14 
using the Tillinghast Reflector Echelle Spectrograph \citep[TRES;][]{furesz08} on the Fred Lawrence 
Whipple Observatory 1.5\,m telescope.  TRES delivers a 51-order echelle spectrum covering the 
full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
standard techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for more 
specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.

\citet{torres12} originally acquired these data in an effort to measure the WASP-14 stellar 
properties, and thereby better constrain the parameters of the transiting exoplanet it hosts.  To 
that end, they employed the {\tt MOOG} and {\tt SPC} approaches outlined in Section 
\ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
free parameter or assumed to be fixed (to $\log g = 4.29$), based on an inference of the mean 
stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
evolution models in the H-R diagram \citep{joshi09}.  Only parameter inferences from the latter 
assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
synthesize a suite of individual \ion{Fe}{1} and \ion{Fe}{2} lines in the $\sim$4300--6750\AA\ 
region, whose equivalent widths were compared with the TRES measurements in a simple least-squares 
sense.  This approach suggests that $T_{\rm eff} = 6150\pm75$\,K and $[{\rm Fe/H}] = 
-0.34\pm0.10$\,dex.  The {\tt SPC} analysis relied on a direct comparison between three TRES 
spectral orders covering $\sim$5100-5400\,\AA\ and synthetic spectra from the {\sc CfA/Kurucz} 
library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
i = 3.9\pm0.5$\,km s$^{-1}$.  It is worth noting that these parameter inferences represent an 
average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
the systematic spread between the best-fit parameter values. \comm{This is something we need to 
discuss more with Dave and Willie.}

Given the decision to employ a spectral library to generate models
(Sect.~\ref{subsec:synthetic}), it is most relevant to compare the
\citeauthor{torres12}~{\tt SPC} result with our framework.  Using the same TRES
spectral orders (for only a single WASP-14 observation, as noted above), the
{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log
g$, we infer parameters that are entirely consistent with the {\tt SPC}
approach: $T_{\rm eff} = 6517\pm13$\,K, $[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and
$v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~\ref{fig:Kurucz_posterior}
displays some relevant marginal projections of the posterior-space, and
Table~\ref{table:Kurucz} compiles our estimates of the peaks of the marginal
posteriors and associated confidence intervals for all parameters employed in
the modeling.  Figure~\ref{fig:Kurucz_residuals} demonstrates the overall
quality of this approach in reproducing the WASP-14 spectrum and, most
significantly, the structure of the associated residuals.
Figure~\ref{fig:metacomparison} compares the inferred parameters with the
various determinations made by \citet{torres12}.

\begin{figure}[!htb]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
  \caption{The posterior probability function of the stellar parameters for WASP-14 as explored by the MCMC Gibbs sampler. These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$). Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
}
\label{fig:Kurucz_posterior}
\end{center}
\end{figure}

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_Kurucz_logg.pdf}
\caption{A sample of the WASP-14 data set overlaid with a model spectrum produced by parameters drawn from the posterior probability distribution (fixing $\log_{10} g = 4.29$ dex) and using the {\sc CfA/Kurucz} synthetic library. The residuals are shown zoomed along with the three-sigma contours of random draws from the covariance matrix, after Figure~\ref{fig:matrix}.}
\label{fig:Kurucz_residuals}
\end{center}
\end{figure*}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6517 & 13 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex\tablenotemark{a}} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.62 & 0.02 $\kms$ \\
$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
$\cc{22}{1}$ & -0.014 & 0.002\\
$\cc{22}{2}$ & -0.014 & 0.002\\
$\cc{22}{3}$ & -0.003 & 0.001\\
\\
$b_{22}$ & 1.04 & 0.02 \\
$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
$l_{22}$ & 7.5 & 0.8 $\kms$ \\

\cutinhead{Order 23 nuisance parameters}
$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
$\cc{23}{1}$ & -0.013 & 0.002 \\
$\cc{23}{2}$ & -0.025 & 0.001 \\
$\cc{23}{3}$ &  0.001 &  0.001\\
\\
$b_{23}$ & 1.01 & 0.02 \\
$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
$l_{23}$ & 6.5 & 0.5 $\kms$\\

\cutinhead{Order 24 nuisance parameters}
$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00\tablenotemark{a}} \\
$\cc{24}{1}$ & -0.010 & 0.002 \\
$\cc{24}{2}$ & -0.011 & 0.001\\
$\cc{24}{3}$ &  0.000 & 0.001\\
\\
$b_{24}$ & 1.02 & 0.02 \\
$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
$l_{24}$ & 7.3 & 0.6 $\kms$ \\
\enddata
\tablenotetext{a}{parameter value fixed}
\end{deluxetable}

@@

To understand the apparent differences in the quoted precisions, it is beneficial to review what factors 
into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
range that no local covariance kernels need to be instantiated; in this specific case we use only a 
global covariance kernel.}  To break down their respective contributions, we performed the 
same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
are generated from the model; in this paradigm, the only way to introduce deviation is through 
white noise (see Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is invalid, since just a few simple parameters 
($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
posteriors. For synthetic models which are less accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend to dominate the error budget.

In addition to statistical and systematic error, one might expect that systematic instrumental error might also contribute to the parameter uncertainties. Commonly, one might estimate this error by fitting spectra taken by different instruments or at different times and then averaging the results. Additionally, one may add in an additional uncertainty ``floor'' that maps to the empirical scatter of typical spectroscopic fits. Such procedures generally make the assumption that this additional error contribution is Gaussian and independent in each parameter of interest. Our hierarchical Bayesian model avoids such a heuristic by
parameterizing the quality of fit self-consistently: because the parameter uncertainty is derived through the model itself, it preserves correlations between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$. Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters. However, for well-calibrated data sets from the same instrument any ill-effects are generally small. To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim10$ K and $\lesssim 0.001$ dex in $\Z$. When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.
